The Emperor Has No Clothes

### 2011-05-02 Winds/papers/LBG_Lya/lya_lbg.tex
jxp
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5949 ac11df5e-a788-c23b-a8e1-c758c323924d

The LBG clump model proposed by S+10 envisions an isotropic outflow of
dense clumps, each optically thick to \lya\ and metal-line
transitions, but which as an ensemble have a partial covering fraction to
the LBG and background sources.  We discuss in PKR11 our radiative
transfer treatment for this wind in the context of \mgiid.  In the
following, we adopt the same algorithms but focus on
\lya\footnote{Following standard practice, we treat \lya\ as a single
  transition instead of a double because the line separation is much less
  than the astrophysical velocities of interest here.}
and modify the parameters of the LBG clump model as appropriate.

In contrast to PKR11, we include two additional features in the
model.  First, we have explored two intrinsic emission models for the
galaxy near the \lya\ line: 
(i) a flat continuum, assumed to be generated by O and B stars within the
star-forming galaxy; and
(ii) a continuum plus \lya\ emission-line model parameterized by a
line-width $\sigma_{\rm Ly\alpha}$ and a \lya\ equivalent width \wlya.
The \lya\ emission is presumed to be generated by \ion{H}{2} regions within the
galaxy.  We emphasize that neither of these simple models are likely
to describe the complex emission of a star-forming galaxy at \lya.
Nevertheless, they help capture some of the salient features of the
LBG clump wind model.

Second, we have modeled the attenuation of photons near \lya\ by the
IGM surrounding LBGs on scales of $\sim 100\,
{\rm kpc} - 10$\,Mpc.  Our approach follows the method of
\cite{santos04}, using the \cite{barkana04} model for infall onto
galactic halos at $z \approx 3$.  We assume a galaxy with dark matter
halo mass $M_{\rm DM} \approx 10^{12} \msun$, virial radius $\mrvir 
= 80$\,kpc, circular velocity $v_{\rm c} = 220 \mkms$, and an
IGM with $T = 10^4$\,K.  THis amterial is bathed in a uniform
extragalactic UV
background (EUVB) radiation field with ionization rate $\Gamma =
10^{-12.5} \, \rm s^{-1}$ \citep{fpl+08}.  With this parameterization,
we can calculate the density and velocity profiles of gas at $r >
\mrvir$ for the LBG (and assume the gas at $r\le \mrvir$ is dominated by
the outflow) and apply the Sobolev approximation to calculate the
optical depth \tigm\ to \lya\ as a function of velocity.
We caution that
this $\exp(-\mtigm)$ prescription does not capture the complexity of
gas on these scales.  Indeed, \cite{zheng10a} have argued that a more
proper treatment tends to yield even higher average opacity.
[INCLUDE THE BOOST? Yes, and justify to match $D_A$]
[CHECK against Djisktra]
Therefore, one may wish to consider our estimation of \tigm\ as a
lower limit.

Figure~\ref{fig:1D} (top panels) presents the 
calculated 1D \lya\ spectrum for the
two intrinsic emission models of the galaxy.
For the line-emission, we have adopted $\sigma_{\rm Ly\alpha} =
50\mkms$ and $\mwlya = 100$\AA, the latter of which is suggested for
actively star-forming galaxies \citep{lya_emit}.
The second panels show the predicted results when one includes
radiative transfer through the LBG clumpy wind model.  Similar to the \mgii\
results of PKR11, we observe significant absorption by the outflow for
$v < 0\mkms$.  
Although the opacity of the outflow peaks at $v \approx 0 \mkms$, the
minimum flux is actually observed to be closer to $v \approx -300
\mkms$ because
scattered emission `fills-in' the absorption at $v \lesssim 0
\mkms$.  Most of these scattered photons were
emitted transverse to our sightline and then scattered once to
Earth.  The tail of emission to positive velocities ($v
\approx +500\mkms$) in contrast, are associated with photons that
scattered off the far-side of this isotropic outflow.  
We emphasize that this emission is weak in comparison with the
flux at $v \approx 0 \mkms$, simply because of geometrical projection.

[Emphasize that in this clump model, photons don't diffuse in
frequency much]

The third pair of panels includes the effects of IGM scattering.  For
the DM halo and IGM model assumed, we find a peak optical depth
$\mtmax \approx 7$ at $v \approx v_{\rm c}$ that rapidly declines to
$\mtauv < 0.5$ at $v=0\mkms$.  This implies complete attenuation of
the \lya\ emission at $v \approx v_{\rm c}$ but only modest
attenuation at lower velocities.  Although the IGM effects for our
parameterization are relatively modest compared to the outflow, it is evident that
the IGM cannot be ignored at this redshift.  If the IGM were just $5
\times$ more neutral than modeled here, the \lya\ line would be
completely ``extinguished'' at $v \gtrsim 0 \mkms$, as predicted for
the $z \sim 6$ universe \citep{zheng+10a}.  By the same token, if the
galaxy emits ionizing photons then \tmax\ may fall below unity and the
IGM would play a minor role.  Clearly, a full treatment 
is warranted for the $z \sim 3$ universe.

In the bottom panels we present the full model, convolved with an
ideal spectrometer having FWHM~$\approx 250\mkms$, compared against the
average \lya\ emission profile observed for a sample of $z \sim 3$ 
LBGs \citep{shapley03}. 
[Match our resolution to Shapley??]
At this spectral resolution, which is
typical of LBG spectroscopic observations, the weak emission at $v
\approx 0 \mkms$ that was
apparent for the continuum model is lost and one predicts a simple
absorption profile that extends to positive velocity. 
This model is a very poor description of the data.  
For the line+continuum intrinsic model, in contrast, it is difficult
to detect the outflow and the IGM attenuation is completely lost at
this spectral resolution within the \lya\ emission profile.
Neither model predicts significant line-emission 
at several hundred \kms\ redward of systemic.  
Such emission is {\it not a natural prediction of a clumpy outflow
  model} where the photons tend to escape because of partial covering
not photon diffusion.  In fact, reviewing the
results on \mgii\ from PKR11, such shifts are rare in
outflow models that we have considered.  The only models in PKR11
with significant shifts to positive velocity was when there was large 
opacity at $v \approx 0 \mkms$ near the source, e.g.\ models that
included an ISM component. In essence, this gas modifies the intrinsic
line profile of \lya, yielding the standard `double hump' emission and
then the blue hump is strongly attenuated by the outflow.
[Test this with Lya!] 
It is also worth noting that a strict infall scenario on Mpc scales
also naturaly predicts a redshifted \lya\ line.  The IGM model of
\cite{zheng+10}, for example, predicts \lya\ emission will be detectd
at $v \approx +XX \mkms$ from star-forming galaxies.

In contrast to our assertions,
other authors have modeled the \lya\ profiles of LBGs and have
concluded the redshifted emission is a `clear signature' of an
outflow.
The [most extensive] modeling of \lya\ emissoin by LBGas to date has
been performed by A. Verhomme and collaborators
\citep{verhommea,verhommeb}.  Their preferred model has been a thin
expanding shell characterized by the expansion speed $v_{\rm exp}$ and
total surface density of \ion{H}{1} gas in the shell.  Although the
shell's dynamics play a role in the \lya\ emission kinematics, the
[dominant] aspect for the emergent spectrum is the \ion{H}{1} column
density.  Indeed, they only predict a dominant emission peak at $v \gg
0 \mkms$ if $\mnhi > 2\sci{20} \cm{-2}$.  For lower values, the
emission is preferentially at $v \lesssim +100 \mkms$.  The
explanation is similar to the ISM model described above;  large \nhi\
values force diffusion of \lya\ in frequency.  Furthermore, their
models predict substantial \lya\ absorption at $v < 0 \mkms$.  
Strong absorption is observed in a handful of individual spectra
\citep[e.g.][]{cb58,dessauges}, but is not evident in the stacked
spectra of \cite{shapley93} or \cite{steidel+10}.  
[Why is there sometimes weak \lya\ emission at $v < 0 \mkms$ in LAEs
at $z \sim 3$?]
[Read Shapley et al 2009 paper]
[Redshfited, asymmetric emission is suggestive of an outflow]

Now consider the spatial distribution of \lya\ emission.
Figure~\ref{fig:SB} shows the predicted surface brightness profile
vs.\ projected radius for the flat-continuum intrinsic model and with
no IGM attenuation.  The signal is integrated over velocities $v =
[-100,400]\mkms$, which corresponds to the majority of predicted line-emission.
Overplotted on the model, at an arbirtrary
normalization, is the fitted profile of S+11 to their observations,
$\mu(\rho) \propto \exp[-\rho/b]$ with $b=20$\,kpc.  
Although this model appears to be tending toward such an exponential
profile at very large impact parameter ($\rho > 50$\,kpc), the model
predict greater flux and a steeper fall-off at small radii.
In strong constrast to
the observations, the clump model predicts that 95\%\ of the
integrated flux is emitted at $\rho < 10$\,kpc.  
It is simply impossible to reconcile the LBG clump model with the
observed profile, and we further emphasize that including the IGM
attenutation and/or an intrinsically bright \lya\ emission line only
worsens the agreement.

The explanation for the model's failure is twofold:  
(1) the covering fraction of the clumps decrease with increasing radius to
$f_{\rm c} < 0.2$ at $r > 50$\,kpc; and
(2) the velocity gradient is very shallow beyond the inner few kpc. Therefore,
the outflow absorbs very few photons at large radius compared to the
inner regions.  In terms of the Sobolev approximation, we estimate
$\tau_{\rm S}(\rho = 60 \, {\rm kpc}) = 10^{-4} \tau_{\rm S}(\rho=1\, {\rm
  kpc})$.  [Figure?]
We further emphasize that the inclusion of an ISM
component at small radii would not qualitatively alter any of these conclusions.
Instead, one requires additional sources of \lya\ emission at large
radii or additional opacity in the CGM far beyond that predicgted by
the LBG clump model.  We return to these issues in
$\S$~\ref{sec:later}.

@@

Both the energy and power of this outflow are predicted to achieve
very large quantities and, again, most of this occurs at large radii.
[Discuss \lya\ emission here]

>>>>

The LBG clump model proposed by S+10 envisions an isotropic outflow of
dense clumps, each optically thick to \lya\ and metal-line
transitions, but which as an ensemble have a partial covering fraction
$f_{\rm c}$ to
the LBG and background sources.  We discuss in PKR11 our radiative
transfer treatment for this wind in the context of \mgiid.  In the
following, we adopt the same algorithms but focus on
\lya\footnote{Following standard practice, we treat \lya\ as a single
  transition instead of a doublet because the line separation $(\delta
  v \approx 3.4 \mkms$) is less
  than the astrophysical velocities of interest here.}
and modify the parameters of the LBG clump model as appropriate.

In contrast to PKR11, we include two additional features in the
model.  First, we have explored two intrinsic emission models for the
galaxy near the \lya\ line: 
(i) a flat continuum, assumed to be generated by O and B stars within the
star-forming galaxy; and
(ii) a continuum plus single \lya\ emission-line model parameterized by a
line-width $\sigma_{\rm Ly\alpha}$ and a \lya\ equivalent width \wlya.
The \lya\ emission is presumed to be generated by \ion{H}{2} regions within the
galaxy.  We emphasize that neither of these simple models are likely
to describe the complex emission of a star-forming galaxy at \lya.
Nevertheless, they reveal the salient features of the
LBG clump wind model.

Second, we have modeled the attenuation of photons near \lya\ by the
IGM surrounding LBGs on scales of $\sim 100\,
{\rm kpc} - 10$\,Mpc.  Our approach follows the method of
\cite{santos04}, using the \cite{barkana04} model for infall onto
galactic halos at $z \approx 3$.  We assume a galaxy with dark matter
halo mass $M_{\rm DM} \approx 10^{12} \msun$, virial radius $\mrvir 
= 80$\,kpc, circular velocity $v_{\rm c} = 220 \mkms$, and an
IGM with $T = 10^4$\,K.  This material is bathed in a uniform
extragalactic UV
background (EUVB) radiation field with ionization rate $\Gamma =
10^{-12.5} \, \rm s^{-1}$ \citep{fpl+08}.  With this parameterization,
we can calculate the density and velocity profiles of gas at $r >
\mrvir$ for the LBG (and assume the gas at $r\le \mrvir$ is dominated by
the outflow) and apply the Sobolev approximation to calculate the
optical depth \tigm\ to \lya\ as a function of velocity.
We have further increased this optical depth by a factor of 2 to
better match the average opacity observed for the IGM at this redshift
\citep[e.g.][]{faucher08}. 
We caution that
this $\exp(-\mtigm)$ prescription does not capture the complexity of
gas on these scales.  Indeed, \cite{zheng10a} have argued that a more
proper treatment tends to yield even higher average opacity.
[CHECK against Djisktra]
Therefore, one could consider our estimation of \tigm\ to be a
lower limit.

Figure~\ref{fig:1D} (top panels) presents the 
calculated 1D \lya\ spectrum for the
two intrinsic emission models of the galaxy.
For the line-emission, we have adopted $\sigma_{\rm Ly\alpha} =
50\mkms$ and $\mwlya = 100$\AA, the latter of which is estimated for
actively star-forming galaxies \citep{lya_emit}.
The second panels show the predicted results when one includes
radiative transfer through the LBG clumpy wind model.  Similar to the \mgii\
results of PKR11, we observe significant absorption by the outflow for
$v < 0\mkms$.  
Although the intrinsic opacity of the outflow peaks at $v \approx 0 \mkms$, the
minimum flux is actually observed to be closer to $v \approx -300
\mkms$ because
scattered emission `fills-in' the absorption at $v \lesssim 0
\mkms$.  Most of these scattered photons were
emitted transverse to our sightline and then scattered once to
Earth.  The tail of emission to positive velocities ($v
\approx +500\mkms$) in contrast, are associated with photons that
scattered off the far-side of this isotropic outflow.  
We emphasize that this emission is weak in comparison with the
flux at $v \approx 0 \mkms$, simply because of geometrical effects.
Clumping models differ qualitatively from diffuse medium scenarios
with high optical depth in that the latter forces the diffusion of
\lya\ in frequency space.  In clump models, especially those with
partial covering, the emission occurs much closer to systemic.

The third pair of panels includes the effects of IGM scattering.  For
the DM halo and IGM model assumed, we find a peak optical depth
$\mtmax \approx 7$ at $v \approx v_{\rm c}$ that rapidly declines to
$\mtauv < 0.5$ at $v=0\mkms$.  This implies complete attenuation of
the \lya\ emission at $v \approx v_{\rm c}$ but only modest
attenuation at lower velocities.  Although the IGM effects for our
parameterization are relatively modest compared to the outflow, it is evident that
the IGM cannot be ignored at this redshift.  If the IGM were just $5
\times$ more neutral than modeled here, the \lya\ line would be
completely ``extinguished'' at $v \gtrsim 0 \mkms$, as predicted for
the $z \sim 6$ universe \citep{zheng+10a}.  By the same token, if the
galaxy emits ionizing photons then \tmax\ may fall below unity and the
IGM would play a very minor role.  Clearly, a full treatment 
is warranted for the $z \sim 3$ universe.

In the bottom panels we present the full model, convolved with an
ideal spectrometer having FWHM~$\approx 250\mkms$, compared against the
average \lya\ emission profile observed for a sample of $z \sim 3$ 
LBGs \citep{shapley03}. 
[Match our resolution to Shapley??]
At this spectral resolution, which is
typical of LBG spectroscopic observations, the weak emission at $v
\approx 0 \mkms$ that was
apparent for the continuum model is lost and one predicts a simple
absorption profile that extends to positive velocity. 
This model is a very poor description of the data.  
For the line+continuum intrinsic model, in contrast, it is difficult
to detect the outflow and the IGM attenuation is completely lost at
this spectral resolution within the \lya\ emission profile.
This model predicts significant line-emission, but not 
at several hundred \kms\ redward of systemic.  
Significant line-emission at $\Delta v \gtrsim +400 \mkms$ is
{\it not a natural prediction of a clumpy outflow
  model because the photons do not diffuse in frequency}.
Another point emphasize by the plots is that the clumpy outflow, on
its own, cannot produce a large equivalent width of \lya\ emission.
In an isotropic outflow, the total equivalent width is conserved
(considerations of dust or anisotropy only tend to reduce the
emission, e.g.\ PKR11).  This explains the relatively weak emission
related to the continuum source model.

[Also comment that infall component can easily be lost in \lya]

[Do an ISM + outflow model;  It might predict too much flux at $v
\approx 0 \mkms$]

%In fact, reviewing the
%results on \mgii\ from PKR11, such shifts are rare in
%outflow models that we have considered.  The only models in PKR11
%with significant shifts to positive velocity was when there was large 
%opacity at $v \approx 0 \mkms$ near the source, e.g.\ models that
%included an ISM component. In essence, this gas modifies the intrinsic
%line profile of \lya, yielding the standard `double hump' emission and
%then the blue hump is strongly attenuated by the outflow.
%[Test this with Lya!]

[It is also worth noting that a strict infall scenario on Mpc scales
also naturaly predicts a redshifted \lya\ line.  The IGM model of
\cite{zheng+10}, for example, predicts \lya\ emission will be detectd
at $v \approx +XX \mkms$ from star-forming galaxies.]

Now consider the predicted patial distribution of \lya\ emission.
Figure~\ref{fig:SB} shows the surface brightness profile
versus projected radius for the flat-continuum intrinsic model and with
no IGM attenuation.  The signal has been integrated over velocities $v =
[-100,400]\mkms$, which correspond to the majority of line-emission
(Figure~\ref{fig:1D}).
Overplotted on the model, at an arbirtrary
normalization, is the fitted profile of S+11 to their observations,
$\mu(\rho) \propto \exp[-\rho/b]$ with $b=20$\,kpc.  
%Although this model appears to be tending toward such an exponential
%profile at very large impact parameter ($\rho > 50$\,kpc), 
In contrast, the model
predicts much greater flux and a steeper fall-off at small radii.
In strong conflict with
the observations, the clump model predicts that 95\%\ of the
integrated flux is emitted at $\rho < 10$\,kpc.  
It is simply impossible to reconcile the LBG clump model with the
observed profile, and we further emphasize that including the IGM
attenutation and/or an intrinsically bright \lya\ emission line only
worsens the agreement.

The explanation for the clumpy model's failure is twofold:  
(1) the covering fraction of the clumps decrease with increasing radius to
$f_{\rm c} < 0.2$ at $r > 50$\,kpc; and
(2) the velocity gradient is very shallow beyond the inner few kpc. Therefore,
the outflow absorbs very few photons at large radius compared to the
inner regions.  In terms of the Sobolev approximation, we estimate
$\tau_{\rm S}(\rho = 60 \, {\rm kpc}) = 10^{-4} \tau_{\rm S}(\rho=1\, {\rm
  kpc})$.  
We further emphasize that the inclusion of an ISM
component at small radii or a central source of \lya\ emission only
worsens the disagreement. 
Instead, one requires additional sources of \lya\ emission at large
radii or additional opacity in the CGM far in excess of that predicted by
the LBG clump model.  We return to these issues in
$\S$~\ref{sec:discussion}.

@@

Both the energy and power of this outflow are predicted to achieve
very large quantities and, again, most of this occurs at large radii.
[Discuss \lya\ emission here]
In contrast to our assertions,

\section{Discussion}
\label{sec:discussion}

The results presented in $\S$~\ref{sec:lya} demonstrate that the LBG
clump model proposed by S+10 cannot (on its own) reproduce the
observed \lya\ emission that is characteristic of LBG galaxies.  The
model predicts neither significantly shifted \lya\ emission nor even a
modest surface brightness of extended emission (Figures~\ref{fig:1D},
\ref{fig:SB}).  Although our analysis focused on this specific clump
model, the main findings may be generalized to essenctially all clumpy
wind models, especially those with a partial covering fraction.  
[READ CLUMP PAPERS AGAIN]
In these scenarios, photons do not diffuse in frequency and one
predicts the majority of emission to occur near systemic with a weak
tail to positive velocities.  One can reduce the emission at $v
\approx 0 \mkms$ by assuming a non-isotropic wind or a dusty medium 
(e.g.\ PKR11), but the flux at $\Delta v \gtrsim +400 \mkms$ would
remain very weak. Regarding the surface brightness, clumpy wind scenarios 
are [challenged] to yield significant extended
emission. This is particularly true for winds that increase in
velocity with radius or that have a declining covering
fraction;  one simply has too few photons to scatter at large
radii.  For a flat continuum source, for example, even a wind spanning
$v = 0 - 500 \mkms$ could only absorb 2\AA\ of equivalent width.  
Spread across over 100~kpc$^2$, this implies an essentially negligible
surface brightness.  The only possible [out] is a model that
[resonates] with the very bright \lya\ emission from the galaxy, 
presumably at $v \approx 0 \mkms$.  This would imply clumps at large
radii with modest velocities, i.e.\ a scenario not driven by an
outflow.\footnote{Simlar arguments hold for the even brighter \lya\
  `blobs'.  Indeed, unless one assumes a bright quasar is embedded
  within these phenomena \citep{hpz+09}, one must include additional
 sources of emission throughout the blob.}.

[Should we use Dan's algorithm to invert the surface brightness
profile into a Sobolev solution?]

Before concluding, we comment on two scenarios that may explain the
observed \lya\ emission for LBGs.  Regarding the \lya\ kinematics, we
have emphasized that the observed velocity shift demands a high
opacity gas to force photon diffusion in frequency space.  Following
\cite{neufeld91}, a static slab of gas with optical depth $\tau_0 >
10^X$ and $b \approx 10 \mkms$ (i.e.\ $\mnhi > 10^{X} \cm{-2}$) would
produce significant line-emission at $\Delta v \gtrsim +400 \mkms$.

In contrast to our assertions,
other authors have modeled the \lya\ profiles of LBGs and have
concluded the redshifted emission is a `clear signature' of an
outflow.
The [most extensive] modeling of \lya\ emissoin by LBGas to date has
been performed by A. Verhomme and collaborators
\citep{verhommea,verhommeb}.  Their preferred model has been a thin
expanding shell characterized by the expansion speed $v_{\rm exp}$ and
total surface density of \ion{H}{1} gas in the shell.  Although the
shell's dynamics play a role in the \lya\ emission kinematics, the
[dominant] aspect for the emergent spectrum is the \ion{H}{1} column
density.  Indeed, they only predict a dominant emission peak at $v \gg
0 \mkms$ if $\mnhi > 2\sci{20} \cm{-2}$.  For lower values, the
emission is preferentially at $v \lesssim +100 \mkms$.  The
explanation is similar to the ISM model described above;  large \nhi\
values force diffusion of \lya\ in frequency.  Furthermore, their
models predict substantial \lya\ absorption at $v < 0 \mkms$.  
Strong absorption is observed in a handful of individual spectra
\citep[e.g.][]{cb58,dessauges}, but is not evident in the stacked
spectra of \cite{shapley93} or \cite{steidel+10}.  
[Why is there sometimes weak \lya\ emission at $v < 0 \mkms$ in LAEs
at $z \sim 3$?]
[Redshfited, asymmetric emission is suggestive of an outflow]
